# Reproducible Research: Peer Assessment 1
========================================================


## Loading and preprocessing the data
```{r echo=TRUE}
setwd("C:/Users/Helga/Documents/gabriel/cursos")
# point to activity.csv
bd=read.csv("activity.csv",header=T)
# convert date
bd$date=as.Date(bd$date,"%Y-%m-%d")
# Prepare for English week days
Sys.setlocale("LC_TIME", "English")
sum_steps<- aggregate(steps~date,bd,sum,na.rm=TRUE)
```

## What is mean total number of steps taken per day?
```{r fig.width=7, fig.height=6}
hist(sum_steps$steps,col="Red",main="Total number of steps taken per day frequencies",xlab="Total number of steps")
```
```{r echo=TRUE}
 meansteps = mean(sum_steps$steps,na.rm=T)
  mediansteps = median(sum_steps$steps,na.rm=T)
```

* The __mean__ total number of steps taken per day is _`r meansteps`_
* The __median__ total number of steps taken per day is _`r mediansteps`_

## What is the average daily activity pattern?
```{r echo=TRUE}
mean_steps_interval <- aggregate(steps~interval,bd,mean,na.rm=TRUE)
```
```{r fig.width=7, fig.height=6}
plot(mean_steps_interval$interval,mean_steps_interval$steps,type="l",xlab="Step interval",ylab="Mean number of steps")
```
```{r echo=TRUE}
maxIdx = which.max(mean_steps_interval$steps)
maxInterval = mean_steps_interval[maxIdx, 'interval']
maxAvgStep = mean_steps_interval[maxIdx, 'steps']
```
* On average across all the days in the dataset,the 5 minute interval "__`r maxInterval`__" contains maximum mean number of steps (_`r maxAvgStep`_ steps)

## Imputing missing values
```{r echo=TRUE}
totalNA = sum(is.na(bd$steps))
```
* The total number of __missing (NA)__ values is _`r totalNA`_

Apply a single imputation method (by considering data missing completely at random - MCAR - i guess): substitution by the mean.
```{r echo=TRUE}
naIdx = which(is.na(bd$steps))
naInterval = bd[naIdx, 3]
fillSteps = sapply(naInterval, function(x) { mean_steps_interval[(mean_steps_interval$interval==x), 2]})

nbd = bd
nbd[naIdx, 'steps'] = fillSteps #fill in missing data

NsumStep = aggregate(steps~date, nbd, sum)
```
```{r fig.width=7, fig.height=6}
hist(NsumStep$steps,col="Red",main="Total number of steps taken per day frequencies",xlab="Total number of steps")
meanStepNew = mean(NsumStep$steps)
medianStepNew = median(NsumStep$steps)
```
* The __mean__ total number of steps taken per day was _`r meansteps`_ and now it is _`r meanStepNew`_ thus being the same while the __median__ total number of steps taken per day was _`r mediansteps`_ and now it is _`r medianStepNew`_ which is not a big difference.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
#create a new factor variable "wd" with two levels "Weekday" and "weekend"
bd['wd'] = factor(sapply(bd$date, function(x){ if (weekdays(x) == "Sunday" | weekdays(x) == "Saturday") { "weekend" } else { "weekday"} }))

#calculate the mean number of steps given time interval and wd
meanwd = aggregate(steps~interval + wd, mean, data=bd)

library(lattice)
xyplot( steps ~ interval | wd, data = meanwd, type="l", layout=c(1,2), xlab="Interval", ylab="Mean number of steps")

```